Highly Available Transactions: Virtues and Limitations

### 2013-05-01 writeup/evaluation.tex
aaron, ali, alan comments

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based proposals such as those used to provide weak
isolation in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failures. (Note,
however, that lock-based protocols \textit{do} offer the benefit of
recency guarantees.) While multi-versioned storage systems allow for a
variety of transactional guarantees, few offer traditional weak
isolation (e.g., non-``tentative update'' schemes) in this context.
The MDCC~\cite{mdcc} protocol offers Read Committed isolation with
Lost Update avoidance but is similarly unavailable due to its reliance
on preventing write conflicts. Chan and Gray's read-only transactions
provide read-only transactions with item-cut isolation, causal
consistency, and transactional atomicity (session PL-2L~\cite{adya})
but are unavailable in the presence of coordinator
failure~\cite{readonly}, similar to read-only and write-only
transactions more recently proposed by Eiger~\cite{eiger}. Causal
Serializability offers a similar model (with unavailable
implementation): causal consistency with a variant of Read Uncommitted
between transactions that write to the same data
item~\cite{raynal-causal}.  Brantner's S3 database~\cite{kraska-s3}
and Bayou~\cite{sessionguarantees} can all provide variants of session
PL-2L with high availability, but none provide this HAT functionality
without substantial modification. Swift~\cite{swift} and bolt-on
causal consistency~\cite{bolton} are closest to providing maximum
sticky HAT semantics. As we have seen, it is possible to implement many
guarantees weaker than serializability---including guarantees that are
achievable with high availability---and still not achieve high
availability.

>>>>

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based proposals such as those used to provide weak
isolation in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failures. (Note,
however, that lock-based protocols \textit{do} offer the benefit of
recency guarantees.) While multi-versioned storage systems allow for a
variety of transactional guarantees, few offer traditional weak
isolation (e.g., non-``tentative update'' schemes) in this context.
The MDCC~\cite{mdcc} protocol offers Read Committed isolation with
Lost Update avoidance but is similarly unavailable due to its reliance
on preventing write conflicts. Chan and Gray's read-only transactions
have item-cut isolation with causal consistency and TA (session
PL-2L~\cite{adya}) but are unavailable in the presence of coordinator
failure and assume serializable update transactions~\cite{readonly};
this is similar to read-only and write-only transactions more recently
proposed by Eiger~\cite{eiger}. Causal Serializability offers a
similar model (with unavailable implementation): causal consistency
with a variant of Read Uncommitted between transactions that write to
the same data item~\cite{raynal-causal}.  Brantner's S3
database~\cite{kraska-s3} and Bayou~\cite{sessionguarantees} can all
provide variants of session PL-2L with high availability, but none
provide this HAT functionality without substantial
modification. Swift~\cite{swift} and bolt-on causal
consistency~\cite{bolton} are closest to providing maximum sticky HAT
semantics. As we have seen, it is possible to implement many
guarantees weaker than serializability---including guarantees that are
achievable with high availability---and still not achieve high
availability.

### 2013-05-01 writeup/hat-analysis.tex
aaron, ali, alan comments

As Brewer states, ``systems and database communities are separate but
overlapping (with distinct vocabulary)''~\cite{brewer-slides}. With
this challenge in mind, when possible, we build on existing properties
and definitions from the database and distributed systems literature,
providing a brief, informal explanation and example for each
guarantee. The database isolation guarantees require particular care,
since different DBMSs often use the same terminology for different
mechanisms and may provide additional guarantees in addition to our
implementation-agnostic definitions.  We draw largely on Atul Adya's
dissertation~\cite{adya} and somewhat on its predecessor work: the
ANSI SQL specification~\cite{ansi-sql} and Berenson et al.'s
subsequent 1995 critique~\cite{ansicritique}. We provide a set of
formal definitions, semantics, proofs and observations in our extended
Technical Report~\cite{hat-tr} and opt for a more informal
presentation here.

@@

In our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times, but the order
in which replicas overwrite versions should be consistent with the total order for
each item. \textit{PL-1} requires that writes to different objects are
ordered consistently across transactions, prohibiting Adya's
phenomenon $G0$, also called ``Dirty Writes''~\cite{adya}. If we
create a graph of transactions with edges from one transaction to
another if the former overwrites the latter's write to the same
object, then, under Read Uncommitted, there should be no cycles in
this graph. Consider the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$.  Read Uncommitted is easily
achieved via applying the same logical timestamp to each update in a
transaction and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

@@

\textbf{Repeatable Read} isolation is a confusing property. For
example, some IBM products use the term for fully serializable
isolation~\cite{hat-hotos}, while Gray~\cite{gray-isolation}, Berenson
et al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all
interpret Repeatable Read as providing serializability for all
operations except for predicate-based reads (and so Repeatable Read is
identical to serializability in a key-value store without
predicate-based access). In models like Adya's PL-2.99, the
\textit{Phantom Problem}, whereby two successive predicate-based reads
return different data~\cite{gray-isolation}, is the only kind of
non-serializable behavior allowed. In this paper, we reserve the use
of ``Repeatable Read'' to refer to these definitions, which we will
soon show are not HAT-compliant.

However, as Berenson et al.~\cite{ansicritique} discuss, the ANSI SQL
specification allows several additional behaviors under ``Repeatable
Read'' isolation.  \textbf{ANSI Repeatable Read} requires that, along
with respecting Read Committed isolation, each transaction will only
read one version of each data item that it did not itself produce
(preventing ``Fuzzy Read,'' or P2). In the example below, $T_3$ must
read $a=1$:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: w_x(2)
\\T_3 &: r_x(1)~r_x(a)
\end{align*}
This is a literal interpretation of the phrase ``Repeatable Read'':
unless a transaction modifies a given data item, the observable value
of the data item should not change during the transaction. By itself,
this property is weak---we can always read $\bot$ for each data
item---and easily achieved in a HAT system by caching the values
read. To prevent confusion, we call this isolation model \textbf{Item
  Cut Isolation}: all the transaction's reads should see values from a
non-changing consistent cut or snapshot over the data items. It is
possible to satisfy Item Cut Isolation with high availability by using
multi-versioning on each replica and ensuring that each of a
transaction's reads return an appropriately chosen version (e.g.,
identified by vector clock).

A stronger achievable property that expands on Item Cut Isolation
requires consistent cuts to span both a transaction's predicate reads
(e.g., \texttt{SELECT WHERE}) as well as its item reads.  We call this
\textbf{Predicate Cut Isolation} and it prohibits Phantoms (but does
not guarantee serializability of item-based reads and
writes). Predicate Cut Isolation is also achievable in HAT systems via
similar multi-versioning as above.

@@

TA can be achieved via multi-versioning. If clients attach vector
clocks to every write (incrementing their own position for each
transaction) and replicas store all writes via multi-versioning, then
clients can safely determine which sets of data items to read; this
approach is adopted by Swift~\cite{swift} and, to a lesser-extent,
bolt-on causal consistency~\cite{bolton}. These systems also provide
causal consistency, which requires sticky availability (in each,
implemented via client-side caching). However, TA is achievable
without stickiness via an alternate algorithm: if servers wait to
reveal writes until they are present on all replicas, clients do not
need to be sticky.

@@

The solution to this conundrum is to give up high availability in
favor of sticky availability. Sticky availability permits three
additional models, which we first define and then prove are
unachievable in a generic highly available system:

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client subsequently executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server that is partitioned from the server that executed $T_1$, then
the client will be unable to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes guarantees. However, if the client remains
sticky with the server that executed $T_1$, then we can disallow this
scenario. Accordingly, read your writes, and, by proxy, causal
consistency and PRAM require stickiness. Read your writes is provided
by default in a sticky system, while the remaining causality and PRAM
guarantees can be accomplished with the algorithms for achieving the
remaining session guarantees.

@@

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
application-level consistency guarantees. It can often execute
commutative and logically monotonic~\cite{calm} operations without the
risk of invalidating application-level integrity constraints. We do
not attempt to sketch the entire space of \textit{application-level}
consistency properties that are achievable in HAT systems (see
Section~\ref{sec:futurework}) but we specifically evaluate TPC-C
transaction semantics under HAT consistency guarantees in
Section~\ref{sec:evaluation}.

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ non-failing replicas.

@@

\subsubsection{Unachievable ACID Isolation}

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation, and Cursor Stability guarantees are all unavailable.
Repeatable Read (defined by Gray, Berenson et al., and Adya) and
One-Copy Serializability need to prevent both Lost Update and Write
Skew. These prevention requirements mean that these models are
inherently unavailable.

@@

As we summarize in Table~\ref{table:hatcompared}, a wide range of
isolation levels are achievable in HAT systems, including
transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve read your
writes guarantees and PRAM and causal consistency. However, many other
prominent models, such as Snapshot Isolation, One-Copy
Serializability, and Strict Serializability cannot be achieved due to
the inability to prevent Lost Update and Write Skew phenomena.

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-1em}
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and the system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs
to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute. Using client-side
caching can alleviate some of these challenges~\cite{bolton, swift},
but then the system becomes sticky high available.

>>>>

As Brewer states, ``systems and database communities are separate but
overlapping (with distinct vocabulary)''~\cite{brewer-slides}. With
this challenge in mind, when possible, we build on existing properties
and definitions from the database and distributed systems literature,
providing a brief, informal explanation and example for each
guarantee. The database isolation guarantees require particular care,
since different DBMSs often use the same terminology for different
mechanisms and may provide additional guarantees in addition to our
implementation-agnostic definitions.  We draw largely on Atul Adya's
dissertation~\cite{adya} and somewhat on its predecessor work: the
ANSI SQL specification~\cite{ansi-sql} and Berenson et al.'s
subsequent 1995 critique~\cite{ansicritique}.

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in our extended Technical Report~\cite{hat-tr}. In
our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times, but the order
in which replicas overwrite versions should be consistent with the total order for
each item. \textit{PL-1} requires that writes to different objects are
ordered consistently across transactions, prohibiting Adya's
phenomenon $G0$, also called ``Dirty Writes''~\cite{adya}. If we
create a graph of transactions with edges from one transaction to
another if the former overwrites the latter's write to the same
object, then, under Read Uncommitted, there should be no cycles in
this graph. Consider the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$.  Read Uncommitted is easily
achieved via timestamping each update in a transaction with the same
logical timestamp (which is unique across transactions; e.g., client
ID and logical clock) and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

@@

Several different properties have been labeled \textbf{Repeatable
  Read} Isolation. As we will show in
Section~\ref{sec:unachievable-acid}, some of these are not achievable
in a HAT system. However, the ANSI standardized
implementation-agnostic definition~\cite{ansi-sql} \textit{is}
achievable and directly captures directly the spirit of the term: if a
transaction reads the same data twice, it sees the same value each
time (preventing ``Fuzzy Read,'' or $P2$). In this paper, to
disambiguate between other definitions of ``Repeatable Read,'' we will
call this property ``cut isolation,'' since each transaction reads
from a non-changing cut, or snapshot, over the data items. If this
property holds over reads from discrete data items, we call it
\textbf{Item Cut Isolation (I-CI)}, and if we also expect a cut over
predicate-based reads (preventing Phantoms~\cite{gray-isolation}, or
Berenson et al.'s $P3/A3$~\cite{ansicritique}), we have the stronger
property of \textbf{Predicate Cut-Isolation}. In the example below,
under both levels of cut isolation, $T_3$ must read $a=1$:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: w_x(2)
\\T_3 &: r_x(1)~r_x(a)
\end{align*}
It is possible to satisfy Item Cut Isolation with high availability by
having clients cache their reads such that the value that they read
never changes unless they overwrite it themselves. The cache can be
cleared at the end of each transaction and can alternatively be
accomplished on (sticky) replicas via multi-versioning. Predicate Cut
Isolation is also achievable in HAT systems via similar caching or
multi-versioning that tracks predicate ranges instead of item-based
reads.

@@

TA can be achieved via multi-versioning. If clients attach vector
clocks to every write (incrementing their own position for each
transaction) and replicas store all writes via multi-versioning, then
clients can safely determine which sets of data items to read; this
approach is adopted by Swift~\cite{swift} and, to a lesser-extent,
bolt-on causal consistency~\cite{bolton}. These systems also provide
causal consistency and are subsequently sticky available (each is
implemented via client-side caching). However, TA is achievable
without stickiness via an alternate algorithm: if servers wait to
reveal writes until they are present on all replicas, clients do not
need to be sticky.

@@

The solution to this conundrum is to give up high availability in
favor of sticky availability. Sticky availability permits three
additional guarantees, which we first define and then prove are
unachievable in a generic highly available system:

@@

Read your writes is not achievable in a highly available
system. Consider a client that executes the following two transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)
\\T_2 &: r_x(a)\vspace{-1em}
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client subsequently executes $T_2$ against the same
(partitioned) server, then it will be able to read its
writes. However, if the network topology shifts and the client can
only contact a different server that is partitioned from the server
that executed $T_1$, then the client will be unable to read its own
writes and the system will have to either stall indefinitely to allow
the client to read her writes (violating transactional availability)
or will have to sacrifice read your writes guarantees. However, if the
client remains sticky with the server that executed $T_1$, then we can
disallow this scenario. Accordingly, read your writes, and, by proxy,
causal consistency and PRAM require stickiness. Read your writes is
provided by default in a sticky system. Causality and PRAM guarantees
can be accomplished with well-known variants~\cite{causalmemory,
  bolton, cops, sessionguarantees, swift} of the prior session
guarantee algorithms we presented earlier.

@@

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
application-level consistency guarantees. It can often execute
commutative and logically monotonic~\cite{calm} operations without the
risk of invalidating application-level integrity constraints. We do
not attempt to sketch the entire space of \textit{application-level}
consistency properties that are achievable in HAT systems (see
Section~\ref{sec:relatedwork}) but we specifically evaluate TPC-C
transaction semantics under HAT consistency guarantees in
Section~\ref{sec:evaluation}.

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ non-failing replicas; this affects availability.

@@

\subsubsection{Unachievable ACID Isolation}
\label{sec:unachievable-acid}

@@

Their need to prevent Lost Update means that Consistent Read, Snapshot
Isolation, and Cursor Stability guarantees are all unavailable.
Repeatable Read (defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability need to prevent both Lost Update and Write Skew. These
prevention requirements mean that these guarantees are inherently
unavailable.

@@

As we summarize in Table~\ref{table:hatcompared}, a wide range of
isolation levels are achievable in HAT systems, including
transactional atomicity, cut isolation, and several session
guarantees. With sticky availability, a system can achieve read your
writes guarantees and PRAM and causal consistency. However, many other
prominent semantics, such as Snapshot Isolation, One-Copy
Serializability, and Strict Serializability cannot be achieved due to
the inability to prevent Lost Update and Write Skew phenomena.

@@

\vspace{.5em}\noindent\textbf{Model Composition} Choosing between
combinations of compatible guarantees requires care. Consider the
following transactions:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\\T_3 &: r_x(a)~r_y(b)\vspace{-1em}
\end{align*}
If we want to guarantee both cut isolation and transactional atomicity
and the system only executes $T_1$, $T_2$, and $T_3$, then $T_3$ needs
to read $a=b=\bot$, $a=b=1$, or $a=b=2$. This means that either the
implementation should frequently return $\bot$ (definitely undesirable
and possibly non-convergent), keep multiple versions of each data item
(necessitating potentially complicated distributed garbage
collection), or use pre-declared read sets to fetch a consistent cut
of keys before each transaction begins to execute (e.g.,
COPS~\cite{cops}). Using client-side caching can alleviate some of
these challenges~\cite{bolton, swift}, but then the system becomes
sticky high available.

### 2013-05-01 writeup/hat-appendix.tex
aaron, ali, alan comments

\begin{definition}[Directly Session-Depends by $S$]
\label{def:sd1}
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_i$ immediately precedes $T_j$ in the session commit order for session $S$.
\end{definition}

\begin{definition}[Directly Session-Depends]
\label{def:sd2}
A transaction $T_j$ directly session-depends on transaction $T_i$ if
$T_j$ directly session-depends on transaction $T_i$ for some session
$S$.
\end{definition}

The dependencies for a history $H$ form a graph called its Directed
Serialization Graph ($DSG(H)$).

@@

\begin{definition}[Item Cut Non-Isolation (N-I-CI)]
\label{def:n-ici}
A history $H$ exhibits phenomenon N-ICI if $DSG(H)$ contains a
transaction $T_i$ such that $T_i$ directly item-depends by $x$ on more
than one other transaction.
\end{definition}

@@

\begin{definition}[Predicate Cut Non-Isolation (N-P-CI)]
A history $H$ exhibits phenomenon N-I-CI if $DSG(H)$ contains a
transaction $T_i$ such that, for two predicate-based reads $p_k$,
$p_j$ in $P$ such that $p_k \subseteq p_j$ and $T_i$ directly
predicate-read-depends by $p_k$ on another transaction $T_l$, but
$T_i$ does not directly predicate-read-depend by $p_j$ on $T_l$.
\end{definition}

@@

\begin{definition}[Non-transactional Atomicity (N-TA)]
A history $H$ exhibits phenomenon N-TA if $DSG(H)$ contains a directed cycle
consisting of exactly one read-dependency edge by $x$ from $T_i$ to
$T_j$ and one anti-dependency edge by $y$ from $T_j$ to $T_i$ and
$T_i$'s corresponding read from $x$ does not precede its corresponding
read from $y$.
\end{definition}

@@

@@

\begin{definition}[Not-Read Your Writes (N-RYW)]
A history $H$ exhibits phenomenon N-RYW if $DSG(H)$ contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$, at least one anti-dependency edge, and the remainder
anti-dependency or write-dependency edges.
\end{definition}

>>>>

\begin{definition}[Session-Depends]
\label{def:sd1}
A transaction $T_j$ session-depends on transaction $T_i$ if $T_i$ and
$T_j$ occur in the same session and $T_i$ precedes $T_j$ in the
session commit order.
\end{definition}

The dependencies for a history $H$ form a graph called its Directed
Serialization Graph ($DSG(H)$). If $T_j$ directly write-depends on
$T_i$ by $x$, we draw $T_i\overset{ww_x}\longrightarrow T_j$. If $T_j$
read-depends on $T_i$ by $x$, we draw
$T_i \overset{rw_x}\longrightarrow T_j$. If $T_j$ directly
anti-depends on transaction $T_j$ by $x$, we draw
$T_i \overset{wr_x}\DashedArrow T_j$. If $T_j$ session-depends on
$T_i$ in session $S$, we draw $T_i \overset{S}\longrightarrow T_j$.

@@

\begin{definition}[Item-Many-Preceders (N-I-CI)]
\label{def:n-ici}
A history $H$ exhibits phenomenon N-ICI if $DSG(H)$ contains a
transaction $T_i$ such that $T_i$ directly item-read-depends by $x$ on more
than one other transaction.
\end{definition}

@@

\begin{definition}[Predicate-Many-Preceders (N-P-CI)]
A history $H$ exhibits phenomenon N-I-CI if, for all predicate-based
reads $r_i(P_i:Vset(P_i))$ and $r_j(P_j:Vset(P_j)$ in $T_k$ such that
the logical ranges of $P_i$ and $P_j$ overlap (call it $P_o$), the set
of transactions that change the matches of $P_o$ for $r_i$ and $r_j$
differ.
\end{definition}

@@

\begin{definition}[Observed Transaction Vanishes (N-TA)]
A history $H$ exhibits phenomenon N-TA if $DSG(H)$ contains a directed cycle
consisting of exactly one read-dependency edge by $x$ from $T_i$ to
$T_j$ and one anti-dependency edge by $y$ from $T_j$ to $T_i$ and
$T_i$'s read from $y$ precedes its read from $x$.
\end{definition}

@@

The following session guarantees are directly adapted from Terry et
al.'s original definitions~\cite{sessionguarantees}:

@@

\begin{definition}[Missing Your Writes (N-RYW)]
A history $H$ exhibits phenomenon N-RYW if $DSG(H)$ contains a directed cycle
consisting of a transitive session-dependency between transactions
$T_j$ and $T_i$, at least one anti-dependency edge, and the remainder
anti-dependency or write-dependency edges.
\end{definition}

### 2013-05-01 writeup/hat-vldb.tex
some graphs and joe comments
eval rewrite
More tweaks
more writing and polishing
more writin
aaron, ali, alan comments

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}\\}

@@

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, amsthm, float}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{xspace}

@@

@@

@@

\begin{abstract}
To minimize network latency and remain online during server failures
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. We
analytically and experimentally quantify the availability and
performance benefits of HATs---often several orders of magnitude over
wide-area networks---as well as their necessary semantic compromises.
\end{abstract}

@@

\input{futurework.tex}

@@

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
%\footnotesize
\scriptsize
\balance

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop, scrextend, MnSymbol}

\newcommand{\pbnote}[1]{{\color{red}{#1 ---P.B.}}}

@@

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, amsthm, float}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\newcommand{\subparagraph}{}

\usepackage{xspace, titlesec}

@@

\newcommand{\figfactor}{1.0}

\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx
\newcommand{\commentt}[1]{{\small\texttt{#1}}}

@@

\titleformat*{\section}{\fontsize{12pt}{12}\selectfont\bfseries}
\titleformat*{\subsection}{\fontsize{10pt}{11}\selectfont\bfseries}
\titleformat*{\subsubsection}{\fontsize{10pt}{11}\itshape}

\titlespacing{\section}{1ex}{2ex}{1ex}

\titlespacing{\subsection}{1ex}{1ex}{0.25ex}
\titlespacing{\subsubsection}{1ex}{1ex}{0.25ex}

\let\footnotesize\small

\newcommand*{\DashedArrow}[1][]{\mathbin{\tikz [baseline=-.25ex,-latex, dashed,#1] \draw [#1] (0pt,0.5ex) -- (2em,0.5ex);}}%

@@

\begin{abstract}
To minimize network latency and remain online during server failures
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak isolation, data consistency, and highly
available systems. We analytically and experimentally quantify the
availability and performance benefits of HATs---often several orders
of magnitude over wide-area networks---as well as their necessary
semantic compromises.
\end{abstract}\vspace{-.5em}

@@

%\input{futurework.tex}

@@

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% vldb_sample.bib is the name of the Bibliography in this case
%\footnotesize
\scriptsize

### 2013-05-01 writeup/model.tex
aaron, ali, alan comments

Traditionally, a system provides {\em availability} if every user that
can contact a correct server eventually receives a response from that
server, even in the presence of arbitrary, indefinitely long network
partitions between servers~\cite{gilbert-cap}.\footnote{Under this
  definition, systems that require a majority of servers to be online
  or that require the database to function correctly with high (but
  not perfect) probability are not available.} As in a standard
distributed database, designated servers might perform operations for
different data items; a server that can handle an operation for a
given data item is called a \textit{replica} for that
item.\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{dynamo,
    pnuts, bigtable, spanner, hstore}, we consider partially
  replicated systems.} While this standard model is useful, there are
several other cases we should consider.

>>>>

Traditionally, a system provides {\em high availability} if every user
that can contact a correct (non-failing) server eventually receives a
response from that server, even in the presence of arbitrary,
indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition, systems
  that require a majority of servers to be online or that require the
  database to function correctly with high (but not perfect)
  probability are not available.} As in a standard distributed
database, designated servers might perform operations for different
data items; a server that can handle an operation for a given data
item is called a \textit{replica} for that item.\footnote{There is a
  further distinction between a \textit{fully replicated} system, in
  which all servers are replicas for all data items, and a
  \textit{partially replicated} system, in which at least one server
  acts as a replica for only a subset of all data items. For
  generality, and, for applicability to many modern ``sharded'' or
  ``partitioned'' data storage systems~\cite{dynamo, pnuts, bigtable,
    spanner, hstore}, we consider partially replicated systems.} While
this standard model is useful, there are several other cases we should
consider.
